---
title: "Biweekly Report 1"
author: "Julia Shoobe"
date: "9/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Read in Dataset
```{r}
> suicide_total_deaths <- read_csv("Downloads/suicide_total_deaths.csv")
```

#Libraries
```{r}
install.packages("tidyverse")
library(tidyverse)
```

#Convert to Long
```{r}
suicide.long <- gather(suicide_total_deaths, Year, Deaths, c(2:28))
#creates a dataset with three columns: country, year, and number of deaths
```

#Descriptive Statistics
```{r}
#1990
summary(suicide_total_deaths$`1990`)
#MIN: 1.27; MEDIAN: 471.00; MEAN: 4074.54; MAX: 237,000.00

#1991
summary(suicide_total_deaths$`1991`)
#MIN: 1.30; MEDIAN: 493.00; MEAN: 4169.90; MAX: 242,000.00

#1992
summary(suicide_total_deaths$`1992`)
#MIN: 1.32; MEDIAN: 507.00; MEAN: 4280.00; MAX; 245,000.00

#1993
summary(suicide_total_deaths$`1993`)
#MIN: 1.34; MEDIAN: 539.00; MEAN: 4417.74; MAX: 247,000.00

#1994
summary(suicide_total_deaths$`1994`)
#MIN: 1.36; MEDIAN: 545.00; MEAN: 4542.97; MAX: 249,000.00

#1995
summary(suicide_total_deaths$`1995`)
#MIN: 1.41; MEDIAN: 572.00; MEAN: 4584.30; MAX: 246,000.00

#1996
summary(suicide_total_deaths$`1996`)
#MIN: 1.44; MEDIAN: 549.00; MEAN: 4564.68, MAX: 241,000.00

#1997
summary(suicide_total_deaths$`1997`)
#MIN: 1.49; MEDIAN: 549.00; MEAN: 4552.60; MAX: 236,000.00

#1998
summary(suicide_total_deaths$`1998`)
#MIN: 1.61; MEDIAN: 577.00; MEAN: 4545.24; MAX: 229,000.00

#1999 
summary(suicide_total_deaths$`1999`)
#MIN: 1.64; MEDIAN: 580.00; MEAN: 4558.31; MAX: 218,000.00

#2000
summary(suicide_total_deaths$`2000`)
#MIN: 1.59; MEDIAN: 580.00; MEAN: 4550.00; MAX: 205,000.00

#2001
summary(suicide_total_deaths$`2001`)
#MIN: 1.67; MEDIAN: 589.00; MEAN: 4510.27; MAX: 205,000.00

#2002
summary(suicide_total_deaths$`2002`)
#MIN: 1.70; MEDIAN: 598.00; MEAN: 4526.00; MAX: 209,000.00

#2003
summary(suicide_total_deaths$`2003`)
#MIN: 1.77; MEDIAN: 600.00; MEAN: 4545.22; MAX: 212,000.00

#2004
summary(suicide_total_deaths$`2004`)
#MIN: 1.82; MEDIAN: 587.00; MEAN: 4547.81; MAX: 215,000.00

#2005
summary(suicide_total_deaths$`2005`)
#MIN: 1.89; MEDIAN: 594.00; MEAN: 4540.00; MAX: 218,000.00

#2006
summary(suicide_total_deaths$`2006`)
#MIN: 1.90; MEDIAN: 605.00; MEAN: 4482.0; MAX: 218,000.00

#2007
summary(suicide_total_deaths$`2007`)
#MIN: 1.93; MEDIAN: 615.00; MEAN: 4433.01; MAX: 217,000.00

#2008
summary(suicide_total_deaths$`2008`)
#MIN: 1.98; MEDIAN: 616.00; MEAN: 4397.22; MAX: 216,000.00

#2009
summary(suicide_total_deaths$`2009`)
#MIN: 2.02; MEDIAN: 615.00; MEAN: 4352.42; MAX: 216,000.00

#2010
summary(suicide_total_deaths$`2010`)
#MIN: 2.07; MEDIAN: 615.00; MEAN: 4327.26; MAX: 217,000.00

#2011
summary(suicide_total_deaths$`2011`)
#MIN: 2.12; MEDIAN: 623.00; MEAN: 4305.19; MAX: 222,000.00

#2012
summary(suicide_total_deaths$`2012`)
#MIN: 2.04; MEDIAN: 630.00; MEAN: 4293.18; MAX: 226,000.00

#2013
summary(suicide_total_deaths$`2013`)
#MIN: 2.03; MEDIAN: 645.00; MEAN: 4284.95; MAX: 227,000.00

#2014
summary(suicide_total_deaths$`2014`)
#MIN: 2.10; MEDIAN: 648.00; MEAN: 4303.20; MAX: 230,000.00

#2015
summary(suicide_total_deaths$`2015`)
#MIN: 2.11; MEDIAN: 651.00; MEAN: 4323.96; MAX: 231,000.00

#2016
summary(suicide_total_deaths$`2016`)
#MIN: 2.13; MEDIAN: 671.00; MEAN: 4344.80; MAX: 230,000.00
```

#For my first biweekly report I found a public dataset from gapminder.org. This dataset shows the suicide rate of 187 countries from the years 1990 to 2016. I read the dataset into R-Studio, created a long dataset (I know with certain commands its easier to have long data, so I wanted to create it from the beginning), then I got the descriptive statistics summary (min/max (range), median, and mean). The gather command to create the long dataset, is a part of the tidyverse package, so I downloaded that, as well. The install.packages just needs to be run once; but the library() needs to be rerun everytime. 
